---
title: "PA1_template.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Project submission
####*loading data
```{r libraries}
library(ggplot2)
library(dplyr)
library(lubridate)
```
```{r reading}
###reading the df
steps <- read.csv("/Users/jomr/Desktop/Coursera/Reproducible Research/activity.csv")
```
####*checking the structure and transforming df
```{r str, cache=TRUE}
str(steps)
head(steps)
#converting steps variable into date format
steps$date <- ymd(steps$date)
#converting into tbl_df
steps <- tbl_df(steps)
totalbyday <- dplyr::summarise(steps, totalsteps=sum(steps, na.rm=TRUE))
```
####*producing a histogram of the number of steps by day; then mean and then median; without missing values
```{r hist, results="asis", cache=TRUE, fig.show = 'asis'}
#grouping by date and summarising
steps <- group_by(steps, date)
totalbyday <- dplyr::summarise(steps, totalsteps=sum(steps, na.rm=TRUE))
histogram1 <- hist(totalbyday$totalsteps, main="Total steps by day", xlab="Number of steps")
```

  
*mean and median
```{r mean, results='asis'}
mean1 <- mean(totalbyday$totalsteps)
median1 <- median(as.numeric(totalbyday$totalsteps))
rm(totalbyday)
print(paste("mean1=", mean1))  


print(paste("median1=", median1))
```


```{r intervals1}
#grouping by interval
part2 <- group_by(steps, interval)
part2 <- dplyr::summarise(part2, intervalmean=mean(steps, na.rm=TRUE))
```
  

####*plotting mean number of steps by interval
```{r plot1}
plot(part2$interval, part2$intervalmean, type="l", main = "Mean number of steps by 5 minute interval", ylab="Mean number of steps", xlab="interval (min)")
```
  
####*choosing the interval with the highest average  
```{r interval1}
highest <- part2[which.max(part2$intervalmean),1]
print(highest[[1]])
```
  
#####imputation
```{r imputation, cache=TRUE}
#getting and summing number of NA's
imputeddf <- steps
#imputing missings with the mean for each interval
c <- as.numeric(which(is.na(imputeddf$steps)))
for (i in c) {
            testo <- subset(imputeddf, interval==as.numeric(imputeddf[i,3]), select=steps)
            imputeddf$steps[[i]] <- mean(testo$steps,na.rm=TRUE)
            rm(testo)
} 
rm(c,i)
```
####*producing histogram of steps by day in imputed df, then getting mean and median
```{r histogram2, cache=TRUE, results="asis", fig.show='asis'}
imputeddf <- group_by(imputeddf, date)
histogramdf2 <- dplyr::summarize(imputeddf, totalsteps=sum(steps))
histogram2 <- hist(histogramdf2$totalsteps, main="distribution of steps by day", xlab="total number of steps in a day")


mean2 <- mean(histogramdf2$totalsteps)
median2 <- median(as.numeric(histogramdf2$totalsteps))  
print(paste("mean2=",mean2))
print(paste("median2=",median2))
```


####*comparing means and medians before and after imputation
```{r comparison, results="asis", warnings=FALSE}
library(knitr)
totalbyday <- dplyr::summarise(steps, totalsteps=sum(steps, na.rm=TRUE))
mean1 <- mean(totalbyday$totalsteps)
median1 <- median(as.numeric(totalbyday$totalsteps))
mean2 <- mean(histogramdf2$totalsteps)
median2 <- median(as.numeric(histogramdf2$totalsteps))
compare <- data.frame(c(mean1, mean2), c(median1, median2), row.names=c("original", "imputed"))
colnames(compare)=c("mean", "median")
knitr::kable(compare, caption="mean and median comparison")
```


####*building side by side histograms for comparison
```{r histcompare}
dev.set(which=2)
par(mfrow=c(1,2))
plot(histogram1, xlab="count of total steps by day", main="Original df")
plot(histogram2, xlab="count of total steps by day", main="Imputed df")
par(mfrow=c(1,1)) #setting original graphing pars
```

```{r weekday}
####adding weekday column to imputeddf
imputeddf <- mutate(imputeddf, day=weekdays(date))
for (i in seq_along(1:length(imputeddf$day)))
if (imputeddf$day[[i]] == "Saturday"){imputeddf$weekend[[i]] <- "weekend"} else 
      if (imputeddf$day[[i]] =="Sunday"){imputeddf$weekend[[i]] <- "weekend"
} else {
            imputeddf$weekend[[i]] <- "weekday"
            }
```

```{r final df}
###making summary data frame
weekdayweekend <- group_by(imputeddf, weekend, interval)
wdwkaveraged <- dplyr::summarise(weekdayweekend, averagesteps=mean(steps))
```
###last plot :)
```{r lastplot}
p <- ggplot(data=wdwkaveraged, aes(x=interval, y=averagesteps))
p+geom_line()+facet_grid(weekend~.)+theme_classic()+ylab("Average steps per interval")+xlab("interval (min)")
```
